Ricci curvature of the rooted phylogenetic subtree-prune-regraft graph

### 2015-02-06 curvature.tex
adjacent tree overlap and adjacent tree max curvature lemmas

@@

A lower bound is also available because we can't do better than distance 1 for all trees unless they move $R$.
There are only $\OhOf{n}$ pairs with trees that move $R$.
By ignoring these trees we get a lower bound of $(q_n -(\OhOf{n}))/q_n$ for $W_{1,n}$.

@@

>>>>

Next, we bound the number of neighbors shared by two adjacent trees.

\begin{lemma}
	\label{lem:shared_neighbors}
Let $T$ and $S$ be rooted $X$-trees such that $\dspr{T,S} = 1$.
Then $\size{N(T) \cap N(S)} \le 6n - 17$.
\end{lemma}

\begin{proof}
	$T$ and $S$ differ by one SPR move that moves a subtree $R$.
	Let $U$ be a tree in $N(T)$ and $N(S)$.
	Then either (1) $T$ and $U$ differ in the location of $R$, or (2) $T$ and $U$ differ in the location of another subtree $P$.
	In the latter case, the facts that $T$ and $S$ differ only in the location of $R$ and $\dspr{T,P} = \dspr{S,P} = 1$, imply that $T|(X \setminus L(P) - S|(X \setminus L(P)$.
	Then leaves $r' \in R$, $p' \in P$, and $q' \in Q$, for some subtree $Q$, form a triple of $T$ and a different triple in $S$.
	This incompatible triple can be resolved in at most $6n - 17$ ways, the maximum of which is reached when $P$, $Q$, and $R$ are themselves a ``triple'' of subtrees.
	By Lemma~\ref{ }\todo{need unique move lemma}, each of the subtrees is assigned to at most $2n-6$ unique moves.
	Moreover, one additional overlapping move also moves one of the subtrees (that of the aunt of the LCA of the three subtrees).
	The number of shared neighbors is thus at most $3(2n-6) + 1 = 6n-17$.
	Note that this bound is tight when, for example, $T$ and $S$ are ladders with a different configuration of 3 leaves at maximum depth.
\end{proof}

@@

A lower bound is also available because we can't do better than distance 1 for all trees but the $\OhOf{n}$ pairs from Lemma~\ref{lem:shared_neighbors}.
By ignoring these trees we get a lower bound of $(q_n -(\OhOf{n}))/q_n$ for $W_{1,n}$.

@@

Next we bound the maximum curvature of two adjacent trees.

\begin{lemma}
	The maximum curvature of the uniform random walk between two adjacent trees is $\frac{6n-17}{3n^2-13n+14}$. \todo{lazy random walk is $\frac{6n-15}{3n^2 - 13n + 14}$}
\end{lemma}
\begin{proof}
Let $T$ and $S$ be two adjacent trees.
The maximum curvature occurs when their neighborhoods have maximum overlap, are of equal size, and all other tree pairs are at distance 1.
By Lemma~\ref{lem:shared_neighbors} the maximum overlap is $6n-17$.
The maximum curvature is thus $1 - \frac{\degree{T} - 6n-17}{\degree{T}} = \frac{6n - 17}{\degree{T}}$. This is maximized when $\degree{T}$ is as small as possible, that is $T,S$ are ladders and $\degree{T} = 3n^2 - 13n + 14$.
\end{proof}
